---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# pingAnalysis

<!-- badges: start -->
[![R-CMD-check](https://github.com/atusy/pingAnalysis/workflows/R-CMD-check/badge.svg)](https://github.com/atusy/pingAnalysis/actions)
<!-- badges: end -->

Analyze ping log.

## Installation

Requires R >= 4.1.

You can install the development version from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("atusy/pingAnalysis")
```

## Examples

- Load the package with:

```{r}
library(pingAnalysis)
```

- Prepare CSV file with three columns without header:
    - timestamp with YMDhms-style
    - IPv4 address with prefix-size
    - Ping timing. `-` indicates timeout.

```csv
20200101000001,192.168.1.1/24,1
20200101000002,192.168.1.1/24,-
20200101000003,192.168.1.2/24,1
20200101000004,192.168.1.1/24,-
20200101000005,192.168.1.2/24,-
```

### 1. Treat all ping timeout as server being out of service.

#### Prepare data

```{r}
log_csv <- tempfile(fileext = ".csv")
log_df <- tibble::tibble(
  timestamp = 20200101000000 + seq(5),
  address = "192.168.1.1/24",
  ping = c("1", "-", "1", "-", "-")
)
readr::write_csv(log_df, log_csv, col_names = FALSE)
```

<details><summary>View data</summary>

```{r, echo=FALSE, collapse=FALSE, comment=NULL}
cat(readr::read_lines(log_csv), sep = "\n")
```

</details>

#### Analyze

```{r}
measure_all_timeout(read_log(log_csv))
```

### 2. Treat ping timeout as server being out of service **if timeout continues `N`-times or more**.

Data inherits from the previous example.

#### Analyze

- `N = 1L` is equivalent to `measure_all_timeout()`

```{r}
measure_timeout(read_log(log_csv), N = 1L)
```

- `N = 2L`

```{r}
measure_timeout(read_log(log_csv), N = 2L)
```

### 3. Treat ping result as server being overloaded if rolling mean of ping is greater than or equal to threshold.

#### Prepare data

```{r}
basetime <- 20200101000000
pings <- c(
  0, 11, 0, 11, 12, 0, 11, 12, 13,
  0, 11, 0, 11, 12, 0, 11, NA, 13
) |> as.character() |> dplyr::na_if("0") |> dplyr::coalesce("-")
N <- length(pings) / 2L
log_df <- tibble::tibble(
  timestamp = rep(basetime + seq(N), 2L),
  address = paste0("192.168.1.", 1L:2L, "/24") |> rep(each = N),
  ping = pings
)
readr::write_csv(log_df, log_csv, col_names = FALSE)
```

<details><summary>View data</summary>

```{r, echo=FALSE, collapse=FALSE, comment=NULL}
cat(readr::read_lines(log_csv), sep = "\n")
```

</details>

#### Analyze

Read log

```{r}
log_df <- read_log(log_csv)
```

Analyze varying parameters:

- m: window size for rolling mean
- t: threshold of ping
- N: counts of continuous errors

```{r}
measure_overload(log_df, m = 1L, t = 10, N = 1L)
measure_overload(log_df, m = 2L, t = 10, N = 1L)
measure_overload(log_df, m = 1L, t = 10, N = 2L)
```

### 4. Treat ping result as **subnet**, or network-switch, being out-of-order if all servers fail to response for `N`-times or more.

#### Prepare data

```{r}
basetime <- 20200101000000
pings <- c(
  # 1  2  3  4  5  6  7  8  9 10
  1, 0, 0, 0, 0, 0, 0, 0, 0, 1,
  1, 1, 0, 0, 0, 0, 1, 1, 0, 1,
  1, 1, 1, 0, 0, 0, 0, 1, 0, 1
)
N <- length(pings) / 3L
log_df <- tibble::tibble(
  timestamp = rep(basetime + seq(N), 3L),
  address = paste0("192.168.1.", 1L:3L, "/24") |> rep(each = N),
  ping = ifelse(pings == 0, "-", as.character(pings))
)

log_df |>
  dplyr::mutate(
    address = stringr::str_replace(address, "^192\\.168\\.1\\.", "192.168.2.")
  ) |>
  dplyr::bind_rows(log_df) |>
  readr::write_csv(log_csv, col_names = FALSE)
```

<details><summary>View data</summary>

```{r, echo=FALSE, collapse=FALSE, comment=NULL}
cat(readr::read_lines(log_csv), sep = "\n")
```

</details>

#### Analyze

Read log.

```{r}
log_df <- read_log(log_csv)
```

Assume log contains all the required IP addresses.

```{r}
address_all <- unique(log_df$address)
address_all
```

Do the analysis varying `N`.

```{r}
measure_subnet_timeout(log_df, N = 1L, address_all = address_all)
measure_subnet_timeout(log_df, N = 2L, address_all = address_all)
```


